---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r, echo = FALSE}
# set a seed to avoid unnecessary diffs
set.seed(9284)
```

# shopr

<!-- badges: start -->
<!-- badges: end -->

Use shopr as gym environment in which you can simulate a shop that faces demand for its products and needs to make replenishment decisions based on forecasts to keep the products in stock and turn a profit. Don't evaluate your forecasts by RMSE but by how much money you can make.

## Installation

You can install the development version of `shopr` from [GitHub](https://github.com/timradtke/shopr) with:

```r
# install.packages("devtools")
devtools::install_github("timradtke/shopr")
```

## Get Started

To start a simulation, first initialize the state. `shopr` does not maintain state in memory via objects but writes changes to tables via partitioned parquet files. Specify a path to a directory that `shopr` should use to create the initial set of tables as well as for the following updates.

```{r}
path <- tempdir()
future_end_date <- as.Date("2025-12-31")

shopr::initialize(
  current_date = as.Date("2024-12-31"),
  history_start_date = as.Date("2024-12-01"),
  future_end_date = future_end_date,
  path = path
)
```

The simulation can be performed as a simple loop of daily iterations. Each day starts with an `update()` which receives incoming shipments and places them into inventory, to then serve demand from inventory and observe sales.

Next, you should can decide how you'd like to forecast and decide which orders to for replenishment of products to make. Submit orders as data frame to `shopr::replenish()` which will ensure they arrive in future iterations depending on products' delivery times.

An iteration ends with recording the realized profit and cash flow---mainly to ease later reporting of the overall simulation results.

```{r}
for (day in 1:31) {
  # `update()` is the core of the simulation, updating inventory and sales
  current_date <- shopr::update(path = path)
  
  # While `shopr` provides internal `forecast()` and `make_order()` functions, 
  # they are intended for examples like the one here. You should create
  # forecasts and orders yourself, and pass them back to `shopr` via `replenish()`.
  df_forecast <- shopr:::forecast(
    current_date = current_date,
    future_end_date = future_end_date,
    path = path
  )
  
  df_order <- shopr:::make_order(
    forecast = df_forecast,
    current_date = current_date,
    path = path
  )
  
  # Submit your orders and they will appear in future iterations via `update()`
  shopr::replenish(order = df_order, current_date = current_date, path = path)
  shopr::record_profit_and_loss(current_date = current_date, path = path)
}
```

At the end of the simulation, gather the daily results into the overall profit and loss and cash flow statement:

```{r}
shopr::report_profit_and_loss(
  from_date = as.Date("2025-01-01"),
  to_date = current_date,
  path = path
)
```

As `shopr` maintains its state in parquet tables, you can close your session, have lunch, come back later, and continue the simulation from where you've left off:

```{r}
for (day in 1:3) { # now simulating the first three days of February
  current_date <- shopr::update(path = path)
  # The shop will eventually run out-of-stock as we are not replenishing...
  shopr::record_profit_and_loss(current_date = current_date, path = path)
}

shopr::report_profit_and_loss(
  from_date = as.Date("2025-02-01"),
  to_date = current_date,
  path = path
)
```

## Table Schema

To maintain state in tables, `shopr` comes with an internal definition of tables and their (`arrow`) schemata.

Use `list_tables()` to list all tables that can be managed by `shopr`:

```{r}
tables <- shopr::list_tables()
print(names(tables))
```

Each of the tables comes with an `arrow` schema and partitioning definition:

```{r}
print(tables[["sales"]])
```

Access the tables directly using the `arrow` package, or via the wrapper functions provided by `shopr`. For example, you can open any of the tables as `arrow` data set by passing its name to `open_table()`:

```{r}
ds_sales <- shopr::open_table(table = "sales", path = path)
print(ds_sales)
```

The data can then be queried and read using `dplyr` which can be more efficient when data sets are larger and the table's partitioning can be exploited, for example by filtering for a specific date:

```{r}
ds_sales |>
  dplyr::filter(date == current_date) |>
  dplyr::collect()
```
